### 2008-10-09 trunk/Papers/rmrs-08/entailment.tex
put proofs back in; played around with baselinestretch

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation which may again serve as a starting point
for research along those lines:

\todo{define ``extension''!!}

\begin{prop}\label{thm:big-one}
  Let $\varphi, \varphi'$ be two {\sc rmrs}s.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$. 
\end{prop}

>>>>

At this point, we cannot provide an efficient algorithm for testing
entailment of \rmrs.  However, we propose the following novel
syntactic characterisation which may again serve as a starting point
for research along those lines. We call an \rmrs\ $\varphi'$ an
\emph{extension} of the \rmrs\ $\varphi$ if $\varphi'$ contains all
the EPs of $\varphi$ and $D(\varphi') \supseteq D(\varphi)$.

\begin{prop}\label{thm:big-one}
  Let $\varphi, \varphi'$ be two {\sc rmrs}s.  Then $\varphi \models
  \varphi'$ iff for every solved form $S$ of $\varphi$, there is a
  solved form $S'$ of $\varphi'$ such that $S$ is an extension of
  $S'$. 
\end{prop}
\begin{proof}[Proof (sketch)]
``$\Leftarrow$'' follows from
Props.~\ref{prop:solved-forms-are-satisfiable} and
\ref{prop:models-satisfy-solved-forms}.

``$\Rightarrow$'': We construct a solved form for
$\varphi'$ by choosing a solved form for $\varphi$ and appropriate
substitutions for mapping the variables of $\varphi$ and $\varphi'$
onto each other, and removing all atoms using variables that don't
occur in $\varphi'$ .  The hard part is the proof that the result is a
solved form of $\varphi'$; this step involves proving that if $\varphi
\models \varphi'$ with the same variable assignments, then all {\sc
  ep}s in $\varphi'$ also occur in $\varphi$.
\end{proof}

### 2008-10-09 trunk/Papers/rmrs-08/rmrs.tex
put proofs back in; played around with baselinestretch

Because solved forms are so restricted, it is now possible to `read
off' at least one model from each solved form.  We don't have the
space to present any proofs here; they proceed mostly along the same
lines as the proofs in \newcite{Duchier00dominanceconstraints}.

\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}

@@

\begin{prop} \label{prop:models-satisfy-solved-forms}
  For every tuple $(\tau,\alpha,g,\sigma)$ that satisfies some \rmrs\
  $\varphi$, there is a solved form $\varphi'$ of $\varphi$ such that
  $(\tau,\alpha,g,\sigma)$ also satisfies $\varphi'$.
\end{prop}

\begin{prop}  \label{prop:finite-solved-forms}
  Every {\sc rmrs} $\varphi$ has only a finite number of solved forms, up to
  renaming of variables.
\end{prop}

>>>>

Because solved forms are so restricted, we can `read off' at least one
model from each solved form.

\begin{prop} \label{prop:solved-forms-are-satisfiable}
  Every {\sc rmrs} in solved form is satisfiable.
\end{prop}
\begin{proof}[Proof (sketch; see also
  \cite{Duchier00dominanceconstraints})] 
  For each {\sc ep}, we choose to label the anchor with the
  constructor $p$ of sufficiently high arity whose existence we
  assumed; we determine the edges between an anchor and its children
  from the uniquely determined $\ARG$ atoms; plugging labels into
  holes is straightforward because no label is dominated by more than
  one label; and then we can fill the spaces between the labels and
  anchors with conjunctions.
\end{proof}

@@

\begin{prop} \label{prop:models-satisfy-solved-forms}
  For every tuple $(\tau,\alpha,g,\sigma)$ that satisfies some \rmrs\
  $\varphi$, there is a solved form $\varphi'$ of $\varphi$ such that
  $(\tau,\alpha,g,\sigma)$ also satisfies $\varphi'$.
\end{prop}
\begin{proof}
  We construct the substitution $s$ from $\alpha$ and $g$.  Then we
  add all dominance atoms that are satisfied by $\alpha$ and restrict
  the $\ARG$ atoms to those child indices that are actually used in
  $\tau$.  The result is in solved form because $\tau$ is a tree; it
  is a solved form of $\varphi$ by construction.
\end{proof}

\begin{prop}  \label{prop:finite-solved-forms}
  Every {\sc rmrs} $\varphi$ has only a finite number of solved forms, up to
  renaming of variables.
\end{prop}
\begin{proof}
  Up to renaming of variables, there is only a finite number of
  substitutions on the node and base variables of $\varphi$.  Let $s$
  be such a substitution.  This fixes the set of {\sc ep}s of any
  solved form of $\varphi$ that is based on $s$ uniquely.  There is
  only a finite set of choices for the subsets $S'$ in condition 2 of
  Def.~\ref{defn:solved-form-of}, and there is only a finite set of
  choices of new dominance atoms that satisfy condition 3.  Therefore,
  the set of solved forms of $\varphi$ is finite.
\end{proof}
